Reinforcement Learning Based Transmitter-Receiver Selection for Distributed MIMO Radars

### 2020-01-17 paper.tex

A Distributed multiple-input and multiple-output (MIMO) radar is a system in which multiple transmitters and receivers are deployed in different locations \cite{Haimovich2008}.
The widely distributed antennas increase spatial diversity which can be utilized for improved target localization, parameter estimation and detecting low-observable targets.
In a MIMO radar, the used waveforms are designed jointly and the waveforms are typically orthogonal.
Therefore, each transmitter-receiver (TX-RX) pair forms an independent channel for illuminating and observing the targets.
Increasing the number of channels provides additional spatial diversity at the cost of larger amount of data to be processed and higher power consumption \cite{Godrich2011, Sun2014, Aittomaeki2018}.
In addition, active transmitters may expose their location which might be undesired in certain applications.
Strategies for TX-RX subset selection are studied to preserve the spatial diversity and reduce the costs at the same time.

@@

Performance in target detection and target localization depend on the SINR level \cite{aittomaki1, aittomaki2, antenna_subset_selection, selection_and_placement}.
Therefore, a reward function for reinforcement learning is proposed that depends on the vector $\vsinrexp$ that consists of all the SINR values $\esinrexp$.
The following reward function is used for the employed MAB learning
\begin{equation}\label{eq:reward_func}
    r(\vasvrx, \vasvtx | \vsinrexp) = \frac{1}{N_S M_S}\sum_{n=1}^N \sum_{m=1}^M  \easvrx \easvtx \esinrexp,
\end{equation}
which is a mean of SINR values for the selected channels on a linear scale.
Furthermore, the reward maximization problem can be formulated as
\begin{equation}\label{eq:obj_func}
    \begin{array}{ll}
                &   \max_{\vasvrx, \vasvtx} r(\vasvrx, \vasvtx | \vsinrexp) \\[7pt]
    \text{s.t.} &   
                \left\{\begin{array}{l}
                    \sum_{n=1}^N \easvrx = N_S \\
                    \sum_{m=1}^M \easvtx = M_S.
                \end{array}\right.
    \end{array}
\end{equation}
It is possible to find an optimal solution for the equation (\ref{eq:obj_func}) if $\vsinrexp$ is known.
However, $\vsinrexp$ is not known since we can only obtain the measurements $\esp$. 
Furthermore, it might not be possible to probe all the channels at the same time.
Therefore, $\vsinrexp$ needs to be estimated from the received signal. 
By probing the different channels, an estimate $\vsinrb$ of the channel SINR values can be formed.
An approximate solution for the objective (\ref{eq:obj_func}) is found when the estimate $\vsinrb$ is used to condition the reward function (\ref{eq:reward_func}).
Different subsets must be explored sufficiently large number of times to reinforce the estimate $\vsinrb$ to identify the subset of TX-RX pairs yielding the highest SINR values.

@@

The MAB problem is a sequential decision-making problem in which an agent needs to decide on an action from several competing actions \cite{rl_intro}.
The word arm originates from a slot machine that has a pull lever, and the pull lever is called an arm.
Pulling an arm of a slot machine gives a reward based on a certain probability distribution.
In the MAB problem, the agent needs to sequentially decide from many arms which arm to pull to maximize the cumulative reward.
The strategy of how the agent chooses the arms is called a policy.

@@

To maximize the cumulative reward, the agent needs to find the arm which gives the highest expected reward.
This means that the agent needs to decide when to explore different arms to possibly identify those with higher expected rewards 
and when to keep pulling the arm with currently known highest expected reward.
An estimate for the expected reward is updated each time when the agent has pulled an arm.
The update equation for each arm can be written as
\begin{equation}\label{eq:update1}
    q_{t+1}(a) = q_{t}(a) + \alpha \left(r_t + q_{t}(a)\right),
\end{equation}
where $r_t$ is the received reward from selecting the arm $a$, $\alpha$ is a step size and $q_{t}(a)$ is called action-value \cite{rl_intro}.
For stationary rewards, $\alpha$ can be set to $t^{-1}$, so that (\ref{eq:update1}) calculates the empirical mean. 
For non-stationary rewards, the parameter $\alpha$ needs to be constant so that old rewards have a lower weight than more recent ones \cite{rl_intro}.

In practice, a policy is realized as an algorithm.
We briefly review five different MAB algorithms which are used in the simulations.
The selected algorithms are widely used to solve MAB problems.
These algorithms are
\begin{enumerate}
    \item $\epsilon$-greedy \cite{rl_intro},
    \item Upper Confidence Bound (UCB1) \cite{rl_intro, DSWUCB},
    \item Kullback Leibler Upper Confidence Bound (KL-UCB) \cite{KL_UCB},
    \item Thompson sampling \cite{Thompson, DThompson} and
    \item Recency-Based Exploration (RBE) \cite{RBE,DRBE}.
\end{enumerate}
An analysis for stationary and non-stationary reward distributions can be found in the references.
Also, other algorithms and more extensive discussion on the multi-armed bandits can be found in \cite{lattimorebandit, rl_intro, mab_survey, intro_to_mab}.

>>>>

A Distributed multiple-input and multiple-output (MIMO) radar is a system in which multiple transmitters and receivers are deployed in different locations \cite{Haimovich2008}.
The widely distributed antennas increase spatial diversity which can be utilized for improved target localization, parameter estimation and detecting low-observable targets.
In a MIMO radar, the used waveforms are designed jointly and the waveforms are typically orthogonal.
Therefore, each transmitter-receiver (TX-RX) pair forms an independent channel for illuminating and observing the targets.
Increasing the number of channels provides additional spatial diversity at the cost of larger amount of data to be processed and higher power consumption \cite{Godrich2011, Sun2014, Aittomaeki2011}.
In addition, active transmitters may expose their location which might be undesired in certain applications.
Strategies for TX-RX subset selection are studied to preserve the spatial diversity and reduce the costs at the same time.

@@

Performance in target detection and target localization depend on the SINR level \cite{aittomaeki2011, Aittomaeki2018, Godrich2011, Sun2014}.
Therefore, a reward function for reinforcement learning is proposed that depends on the vector $\vsinrexp$ that consists of all the SINR values $\esinrexp$.
The following reward function is used for the employed MAB learning
\begin{equation}\label{eq:reward_func}
    r(\vasvrx, \vasvtx | \vsinrexp) = \frac{1}{N_S M_S}\sum_{n=1}^N \sum_{m=1}^M  \easvrx \easvtx \esinrexp,
\end{equation}
which is a mean of SINR values for the selected channels on a linear scale.
Furthermore, the reward maximization problem can be formulated as
\begin{equation}\label{eq:obj_func}
    \begin{array}{ll}
                &   \max_{\vasvrx, \vasvtx} r(\vasvrx, \vasvtx | \vsinrexp) \\[7pt]
    \text{s.t.} &   
                \left\{\begin{array}{l}
                    \sum_{n=1}^N \easvrx = N_S \\
                    \sum_{m=1}^M \easvtx = M_S.
                \end{array}\right.
    \end{array}
\end{equation}
It is possible to find an optimal solution for the equation (\ref{eq:obj_func}) if $\vsinrexp$ is known.
However, $\vsinrexp$ is not known since we can only obtain the measurements $\esp$. 
Furthermore, it might not be possible to probe all the channels at the same time.
Therefore, $\vsinrexp$ needs to be estimated from the received signal. 
By probing the different channels, an estimate $\vsinrb$ of the channel SINR values can be formed.
An approximate solution for the objective (\ref{eq:obj_func}) is found when the estimate $\vsinrb$ is used to condition the reward function (\ref{eq:reward_func}).
Different subsets must be explored sufficiently large number of times to reinforce the estimate $\vsinrb$ to identify the subset of TX-RX pairs yielding the highest SINR values.

@@

The MAB problem is a sequential decision-making problem in which an agent needs to decide on an action from several competing actions \cite{Sutton2018}.
The word arm originates from a slot machine that has a pull lever, and the pull lever is called an arm.
Pulling an arm of a slot machine gives a reward based on a certain probability distribution.
In the MAB problem, the agent needs to sequentially decide from many arms which arm to pull to maximize the cumulative reward.
The strategy of how the agent chooses the arms is called a policy.

@@

To maximize the cumulative reward, the agent needs to find the arm which gives the highest expected reward.
This means that the agent needs to decide when to explore different arms to possibly identify those with higher expected rewards 
and when to keep pulling the arm with currently known highest expected reward.
An estimate for the expected reward is updated each time when the agent has pulled an arm.
The update equation for each arm can be written as
\begin{equation}\label{eq:update1}
    q_{t+1}(a) = q_{t}(a) + \alpha \left(r_t + q_{t}(a)\right),
\end{equation}
where $r_t$ is the received reward from selecting the arm $a$, $\alpha$ is a step size and $q_{t}(a)$ is called action-value \cite{Sutton2018}.
For stationary rewards, $\alpha$ can be set to $t^{-1}$, so that (\ref{eq:update1}) calculates the empirical mean. 
For non-stationary rewards, the parameter $\alpha$ needs to be constant so that old rewards have a lower weight than more recent ones \cite{Sutton2018}.

In practice, a policy is realized as an algorithm.
We briefly review five different MAB algorithms which are used in the simulations.
The selected algorithms are widely used to solve MAB problems.
These algorithms are
\begin{enumerate}
    \item $\epsilon$-greedy \cite{Sutton2018},
    \item Upper Confidence Bound (UCB1) \cite{Sutton2018, Garivier2008},
    \item Kullback Leibler Upper Confidence Bound (KL-UCB) \cite{KL_UCB},
    \item Thompson sampling \cite{Thompson, DThompson} and
    \item Recency-Based Exploration (RBE) \cite{RBE,DRBE}.
\end{enumerate}
An analysis for stationary and non-stationary reward distributions can be found in the references.
Also, other algorithms and more extensive discussion on the multi-armed bandits can be found in \cite{lattimorebandit, rl_intro, mab_survey, intro_to_mab}.
